---
title: "Creating clean Nativity Dataset"
author: "Marieke Jones, PhD"
date: "12/10/2019"
output: html_document
---

Load libraries

```{r}
library(tidyverse)
library(forcats)
```

The file I want to load is a fixed width file where the widths are specified by a 95 page PDF table. Using the PDF_Peeler.Rmd file we parsed the PDF table into a useable csv file.

2020-06-03
**There were some issues in the codebook decoder that I fixed by hand** I plan to come back in and fix them programmatically

PDF Peeler should NOT filter out the FILLER columns -- that messes everything up
Codebook decoder has several errors stemming from errors in the codebook provided by the CDC:
- position 1-8 has length 8 not 6
- position 107 has length 1 not 2
- position 365-382 has length 18 not 17

I changed these manually, but I could programmatically change them in the columns dataset

From:
https://www.cdc.gov/nchs/data_access/Vitalstatsonline.htm

Get column information.

>Note that before loading the codebook_decoder, I added a column called Want to the codebook_decoder to mark which columns are of interest

```{r}
columns <- read_csv("codebook_decoder.csv")
column_lengths <- columns %>% pull(Length)
column_names <- columns %>% pull(Variable)
```

Read in full dataset using column widths

```{r}
col_widths <- fwf_widths(column_lengths, column_names)
full_data <- read_fwf("Nat2018.txt", col_positions = col_widths)
```

# HSL workshops

Sample 20000 rows

```{r}
set.seed(6274353) # marieke in phone digits
nat <- sample_n(full_data, 20000)
# rm(nat)
```

Subset for columns of interest

```{r}
keep_cols <- columns %>%
  filter(Want == "X") %>%
  pull(Variable)

nat <- nat %>%
  select(one_of(keep_cols))
```

Resulting Dataset has dimensions: `r dim(nat)`

How many c-sects in subset?

```{r}
nat %>%
  count(DMETH_REC)
```

Recode missings as NA
```{r}
nat <- nat %>%
  mutate(DOB_TT = na_if(DOB_TT, 9999),
         DMAR = na_if(DMAR, 9),
         MEDUC = na_if(MEDUC, 9), 
         FAGECOMB = na_if(FAGECOMB, 99),
         FRACE6 = na_if(FRACE6, 9),
         FEDUC = na_if(FEDUC, 9),
         PRIORLIVE = na_if(PRIORLIVE, 99),
         PRIORTERM = na_if(PRIORTERM, 99),
         TBO_REC = na_if(TBO_REC, 9),
         PREVIS = na_if(PREVIS, 99),
         M_Ht_In = na_if(M_Ht_In, 99),
         BMI = na_if(BMI, 99.9),
         DWgt_R = na_if(DWgt_R, 999),
         WTGAIN = na_if(WTGAIN, 99),
         RF_CESARN = na_if(RF_CESARN, 99),
         NO_RISKS = na_if(NO_RISKS, 9),
         ME_PRES = na_if(ME_PRES, 9),
         ME_ROUT = na_if(ME_ROUT, 9),
         RDMETH_RE = na_if(RDMETH_RE, 9),
         DMETH_REC = na_if(DMETH_REC, 9),
         ATTEND = na_if(ATTEND, 9),
         APGAR5 = na_if(APGAR5, 99),
         COMBGEST = na_if(COMBGEST, 99),
         DBWT = na_if(DBWT, 9999))
```

Drop a few more variables

```{r}
nat <- nat %>%
  select(-MAR_P, -RF_CESAR, -RDMETH_RE)
```

Use forcats to recode variables
fct_recode(x, new = "old", new = "old")

BFACIL3:
1 In Hospital
2 Not in Hospital
3 Unknown or Not Stated

MRACE6:
1 White (only)
2 Black (only)
3 AIAN (only)
4 Asian (only)
5 NHOPI (only)
6 More than one race

DMAR:
1 Yes
2 Unmarried

MEDUC:
1 8th grade or less
2 9th through 12th grade with no diploma
3 High school graduate or GED completed
4 Some college credit, but not a degree.
5 Associate degree (AA,AS)
6 Bachelor’s degree (BA, AB, BS)
7 Master’s degree (MA, MS, MEng, MEd, MSW, MBA)
8 Doctorate (PhD, EdD) or Professional Degree (MD, DDS,
DVM, LLB, JD) 
9 Unknown

FRACE6
1 White (only)
2 Black (only)
3 AIAN (only)
4 Asian (only)
5 NHOPI (only)
6 More than one race
9 Unknown or Not Stated

ME_PRES
1 Cephalic
2 Breech
3 Other
9 Unknown or not stated

ME_ROUT
1 Spontaneous 2 Forceps
3 Vacuum
4 Cesarean
9 Unknown or not stated

DMETH_REC
1 Vaginal
2 C-Section 
9 Unknown

ATTEND
1 Doctor of Medicine (MD)
2 Doctor of Osteopathy (DO)
3 Certified Nurse Midwife (CNM) 4 Other Midwife
5 Other
9 Unknown or not stated

```{r}
nat <- nat %>%
  mutate(BFACIL3 = fct_recode(as.factor(BFACIL3), 
                              Hosp = "1", 
                              NoHosp = "2", 
                              Unknown = "3"),
         MRACE6 = fct_recode(as.factor(MRACE6), 
                             White = "1", 
                             Black = "2", 
                             AIAN = "3", 
                             Asian = "4", 
                             NHOPI = "5", 
                             MoreThan1 = "6"),
         DMAR = fct_collapse(as.factor(DMAR), 
                             Married = "1", 
                             Unmarried = "2"),
         MEDUC = fct_collapse(as.factor(MEDUC), 
                              NoHighSchool = "1", 
                              SomeHighSchool = "2", 
                              HighSchoolGrad = "3", 
                              SomeCollege = "4", 
                              Associates = "5", 
                              Bachelors = "6", 
                              Masters = "7", 
                              DoctorateProfess = "8"),
         FRACE6 = fct_recode(as.factor(FRACE6), 
                             White = "1", 
                             Black = "2", 
                             AIAN = "3", 
                             Asian = "4", 
                             NHOPI = "5", 
                             MoreThan1 = "6"),
         ME_PRES = fct_recode(as.factor(ME_PRES), 
                              Cephalic = "1", 
                              Breech = "2", 
                              Other = "3"),
         ME_ROUT = fct_recode(as.factor(ME_ROUT), 
                              Spontaneous = "1", 
                              Forceps = "2", 
                              Vacuum = "3",
                              Cesarean = "4"),
         DMETH_REC = fct_recode(as.factor(DMETH_REC), 
                              Vaginal = "1", 
                              Cesarean = "2"),
         ATTEND = fct_recode(as.factor(ATTEND), 
                              MD = "1", 
                              DO = "2",
                             CNM = "3",
                             OtherMidwife = "4",
                             Other = "5")
         )
```

Check on progress
```{r}
glimpse(nat)
```

The following variables need to be mutated as.numeric
DOB_MM 
DOB_TT
PRIORLIVE
PRIORTERM
PREVIS
WTGAIN
RF_CESARN
APGAR5
DBWT

```{r}
nat <- nat %>%
  mutate(DOB_MM = as.numeric(DOB_MM),
         DOB_TT = as.numeric(DOB_TT),
         PRIORLIVE = as.numeric(PRIORLIVE),
         PRIORTERM = as.numeric(PRIORTERM),
         PREVIS = as.numeric(PREVIS),
         WTGAIN = as.numeric(WTGAIN),
         RF_CESARN = as.numeric(RF_CESARN),
         APGAR5 = as.numeric(APGAR5),
         DBWT = as.numeric(DBWT))
```

We will still need to deal with Y/N questions to set them as 1/0 but I'll have the students do that.

Write out clean datafile
```{r}
write_csv(nat, "Nativity2018_clean.csv")
```
