---
title: "PDF peeler"
author: "David Martin"
date: "12/6/2019"
output: html_document
---

# Goal

We want to load a fixed width format (FWF) file from the CDC for a workshop, specifically from
https://www.cdc.gov/nchs/data_access/Vitalstatsonline.htm. In order to load a FWF file, we will need to specify the widths of each column or the positions of columns of interest.

Along with the FWF datafile, the CDC has published a PDF with a table detailing each column with its width and position.

This script reads the PDF user guide and parses the width and position of each column.

# Load packages

```{r}
# install.packages("pdftools")
library(pdftools)
library(tidyverse)
library(stringr)
```

# Read in the PDF

```{r}
userguide <- pdf_text("UserGuide2018-508.pdf")
```

Split the PDF on the new line `\n`

```{r}
table_data <- userguide %>%
  str_split(pattern = "\n") %>% 
  unlist()
```

Pull out any lines that start with whitespace or with digits since those are lines we are interested in. Set any other lines to "drop"

```{r}
Position <- ifelse(str_sub(str_sub(str_trim(table_data), 1, 50), 1, 1) != " " &
                     str_detect(str_sub(str_sub(str_trim(table_data), 1, 50), 1, 1), "[:digit:]"), 
                   str_trim(str_sub(table_data, 1, 45)), "drop")
```

Look at ones that are not drop

```{r}
as.data.frame(Position) %>%
  filter(Position != "drop")
```

First two are not data lines so drop those along with all the ones called drop
```{r}
pos <- as.data.frame(Position) %>% 
  filter(Position != "1900-2018" & 
           Position != "drop" & 
           Position != "2018 Natality Public Use File") 
```

Main table starts at line 14, so drop lines 1-13

```{r}
pos <- as.data.frame(pos[-c(1:13), ])
```

Main table stops at line 880 and appendix tables are after that so drop all appendices


```{r}
pos <- pos[-c(881:nrow(pos)),]
```

Now separate the entries based on whitespace where the entry has 1 or more characters

```{r}
# split the first row based on whitespace
str_split(pos[1], pattern = " ")
# we get lots of whitespaces in between the entries

# split where the nchar >= 1
new_data <- lapply(str_split(pos, pattern = " "), function(x) x[nchar(x) >= 1])

# keep entries where length is > 0 and < 4
new_data <- new_data[lapply(new_data, length)>0 & lapply(new_data, length) <= 4]
```

Rows should each have 3 entries. Find ones that don't

```{r}
lengths(new_data)
which(lengths(new_data) == 4)
# 19 116 178 197 256 281

new_data[[19]] <- new_data[[19]][-4]
new_data[[116]] <- new_data[[116]][-4] # this is a header
new_data[[178]] <- new_data[[178]][-4] # this is a header
new_data[[197]] <- new_data[[197]][-4] # name got cut
new_data[[197]][3] <- "F_MM_PLAC"
new_data[[256]] <- new_data[[256]][-4]
new_data[[281]] <- new_data[[281]][-4]
```

Turn the output into a dataframe

```{r}
df <- data.frame(matrix(unlist(new_data), ncol = 3, byrow = TRUE))
names(df) <- c("Position", "Length", "Variable")
```

A few entries are not a data row but are a header. We can remove those based on long lengths. This will also get rid of some of the filler columns but that is fine.

```{r}
df <- df %>% 
  mutate(Length = as.numeric(as.character(Length))) %>% 
  filter(Length < 10)
```

Write out the df into a csv that we can use to read in the main file

```{r}
write_csv(df, "codebook_decoder.csv")
```

